# How else can I communicate with APIs from R?

**Learning objectives:**

-   Fetch data from GraphQL APIs.
-   Fetch data from websocket APIs.
-   Recognize gRPC APIs (maybe fetch with httr2? Does that work?)

```{r req-other-setup}
library(ghql)
library(jsonlite)
library(dplyr)
library(httr2)
library(websocket)
library(RProtoBuf)
```

## GraphQL {-}

-   Current landscape feels like "I use GraphQL in other languages, and want it to technically be possible to do so from R."
-   Queries: [ghql](https://docs.ropensci.org/ghql/index.html)
    -   Demo on next slide
-   Server side: [gqlr](http://schloerke.com/gqlr/index.html) (more in later chapter)

## GraphQL setup {-}

```{r req-other-graphql-setup, eval = FALSE}
ghql_con <- GraphqlClient$new(
  url = "https://api.github.com/graphql",
  headers = list(Authorization = paste0("Bearer ", Sys.getenv("GITHUB_PAT")))
)
ghql_con$load_schema()
qry <- Query$new()
```

## GraphQL query {-}

```{r req-other-graphql-query, eval = FALSE}
qry$query('mydata', '{
  repositoryOwner(login:"jonthegeek") {
    repositories(first: 5, orderBy: {field:PUSHED_AT,direction:DESC}, isFork:false) {
      edges {
        node {
          name
          stargazers {
            totalCount
          }
        }
      }
    }
  }
}')
```

## GraphQL response {-}

```{r req-other-graphql-response, eval = FALSE}
x <- ghql_con$exec(qry$queries$mydata) # We created a query named "mydata"
jsonlite::fromJSON(x)
#> $data
#> $data$repositoryOwner
#> $data$repositoryOwner$repositories
#> $data$repositoryOwner$repositories$edges
#>                   node.name node.totalCount
#> 1                     wapir              17
#> 2                    zoomer               3
#> 3                rstats.fyi               0
#> 4                 tidyslack               0
#> 5 jonthegeek.r-universe.dev               0
```

## websocket review {-}

-   Alternative to HTTP
-   `ws://` or `wss://`
-   2-way communication
-   {websocket} package

## Aside: shiny & websockets {-}

-   You might see websocket-related errors in Shiny
-   Shiny UI & server communicate via a websocket connection
-   More directly managed via {httpuv} & TypeScript code
-   Beyond the scope of this book

## websocket demo: setup {-}

```{r req-other-websocket-setup, eval = FALSE}
ws <- WebSocket$new("ws://echo.websocket.events/", autoConnect = FALSE)
ws$onMessage(\(event) {
  now <- format(Sys.time(), digits = 0)
  cat("Client got msg:", event$data, "at", now, "\n")
})
```

## websocket demo {-}

```{r req-other-websocket-demo, eval = FALSE}
ws$connect()
#> Client got msg: echo.websocket.events sponsored by Lob.com at 2023-11-15 2023-11-15 08:28:04
# (can do other things in console now)
1
#> [1] 1
ws$send("hello")
#> Client got msg: hello at 2023-11-15 08:28:10
ws$close()
```

## websocket: toward usefulness {-}

```{r req-other-websocket-setup2, eval = FALSE}
ws_counter <- 1
ws2 <- WebSocket$new("ws://echo.websocket.events/", autoConnect = FALSE)
ws2$onMessage(\(event) {
  ws_counter <<- ws_counter + 1 # Add 1 to global ws_counter var 
  cat(ws_counter, "\n")
})
ws_counter
#> 1
ws2$connect()
#> 2
ws2$send("update")
#> 3
ws2$send("update again")
#> 4
ws2$close()
```

## websocket usecases {-}

-   News/Message feeds
    -   Display new content as it comes in
-   Messaging
    -   Send and receive ~simultaneously without new connections
-   Multi-player games
-   Collaborative editing
-   Real-time dashboards

## gRPC review {-}

-   Google Remote Procedure Call
-   Becoming very popular
-   Good for real-time, 2-way communication
-   Uses HTTP/2 (more socket-like)

## gRPC: What is it really? {-}

-   HTTP/2 (upgraded HTTP)
    -   Can access via {httr2}
-   "Protocol buffers" datatype
    -   Implemented in {RProtoBuf} package

## gRPC demo {-}

From [RProtoBuf paper](https://www.jstatsoft.org/article/view/v071i02)

```{r req-other-grpc}
resp <- request("https://demo.ocpu.io/MASS/data/Animals/pb") |> 
  req_perform()
output <- resp_body_raw(resp) |> 
  unserialize_pb() # This is the important part
identical(output, MASS::Animals)
head(output)
```

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
