# How can I access APIs from R?

**Learning objectives:**

-   Describe how {httr2} calls are structured.
-   Create and modify an {httr2} request.
-   Access an HTTP API with {httr2}.

## What is httr2? {-}

-   Pipe-based API calls
-   *Here we're working with dev (will be released ~soon as 1.0.0)*
-   `request |> perform |> resp (parse)`

## httr2 demo {-}

```{r 03-httr2-demo, eval = FALSE}
candidates <- httr2::request("https://api.open.fec.gov/v1") |> 
  httr2::req_headers("X-Api-Key" = "DEMO_KEY") |> 
  httr2::req_url_path_append("candidates") |> 
  httr2::req_url_query(office = c("H", "S"), election_year = 2024, .multi = "explode") |> 
  httr2::req_perform() |> 
  httr2::resp_body_json()
candidates$results[[1]]$candidate_id
#> [1] "H4OR05312"
```

## URL Pieces {-}

https://api.open.fec.gov/v1/candidates?office=H&office=S&election_year=2024

-   base URL = `https://api.open.fec.gov/v1/`
-   path = `/v1/candidates`
-   query = `?office=H&office=S&election_year=2024`

## Request: URL {-}

-   `request()` set base URL
-   `req_url_path_append()` add parts separated by `/`
-   `req_url_query()` add query parts

```{r 03-req-url-build}
req <- httr2::request("https://api.open.fec.gov/v1") |> 
  httr2::req_url_path_append("candidates") |> 
  httr2::req_url_query(office = c("H", "S"), election_year = 2024, .multi = "explode")
req$url
```

## Request: URL (other functions) {-}

-   `req_url()` *replace* entire URL (don't use this)
-   `req_url_path()` *replace* part after `/` (don't use this)
-   `req_template()` build request with variables

```{r 03-req-template}
req <- httr2::request("https://api.open.fec.gov/v1") |> 
  httr2::req_template("GET /{endpoint}", endpoint = "candidates")
req$url
```

## Request: Headers {-}

We'll see more of this in chapter 5 (authentication)!

```{r 03-headers}
httr2::request("https://api.open.fec.gov/v1/candidates?office=H&office=S") |> 
  httr2::req_headers("X-Api-Key" = "DEMO_KEY", .redact = "X-Api-Key") |> 
  httr2::req_dry_run()
```

## Request: Method {-}

-   `GET` without body
-   `POST` with body
-   `req_method()` to specify (rarely needed)

## Request: Proxy {-}

-   `req_proxy()` to specify proxy
-   Can be useful to see response as if you're somewhere else

## More httr2 to come! {-}

Exact order might change, but:

-   Chapter 4 = Authentication
-   Chapter 5 = Parsing responses
-   Chapter 6 = Pagination (and other strategies for big responses)

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
