# How do I tell the API who I am?

**Learning objectives:**

-   Compare and contrast authentication schemes.
-   Describe the dangers associated with "leaking" credentials.
-   Find authentication information in API docs.
-   Authenticate a request with a username and password.
-   Authenticate a request with an API key.
-   Authenticate a request with the "OAuth dance."
-   Re-authenticate OAuth requests.
-   TODO: LOs about other schemes, cookies?

## But first: Practice safe git {-}

-   Run `usethis::git_vaccinate()`
-   Usually ignore project-level `.Renviron`

## Credential dangers {-}

-   How much account control does it have?
    -   Can it lock you out?
    -   Can it reveal sensitive data?
-   How long does it last?
    -   Shorter = safer
    -   Can you revoke it?
-   How likely is it to leak?
    -   When do you send it?
    -   Where do you send it?
    -   How often do you send it?

## Authentication schemes overview {-}

-   **HTTP Basic Authentication:** username + password sent with request
-   **API Keys/Bearer Tokens:** password-like thing sent with request
    -   Dirty secret: This is what more complex schemes are, eventually
    -   `httr2::req_auth_bearer_token()`
    -   Sometimes: `httr2::req_headers()`
-   **OAuth:** Multi-step process to generate a key
    -   Goal: Make the key as short-lived as possible
    -   Not everyone implements the same way (ie often wrong)
    -   `httr2::req_oauth_*()`
-   **Other:**
    -   Cookies
    -   [Certificates](https://fosstodon.org/deck/@jaredlander/111422477967658857)

## Dangers of leaking credentials {-}

-   游댮 Username + password
    -   Access entire account
    -   Often difficult to turn off/recover
-   游 API key
    -   Can access everything you can access
    -   Often short-lived
    -   You can often revoke this
-   游리 OAuth exists to minimize these dangers

## What does this API want? {-}

-   Find "OpenAPI" or "Swagger" links (or "API json", "API yaml", etc)
    -   Search for "secrutitySchemes"
    -   Often easier-to-understand details than docs!
-   Might need to do something to "register"
    -   "Request an API key"
    -   "Register your App" (or "Client") (see OAuth)

## HTTP Basic Authentication {-} 

-   `httr2::req_auth_basic(req, username, password = NULL)`
-   Leave password blank: Prompt interactive semi-securely
-   Avoid using this if you can!
    -   I can't remember any API that only offers this

## API Keys & Bearer Tokens {-}

-   `httr2::req_auth_bearer_token(req, token)` (specific header)
    -   Pass `token` as `Sys.get("API_TOKEN_NAME")`
    -   Save token with `usethis::edit_r_environ()`
-   Catch-all: `httr2::req_headers(.req, ..., .redact = NULL)`
    -   `...` = `token_parameter = Sys.get("API_TOKEN_NAME")`
    -   `.redact = "token_parameter"` to hide in print

## OAuth: Terminology {-}

-   **User:** You, or someone using your code
-   **Client:** The thing asking for permission to act as the user
    -   Your code, but also
    -   A reusable thing that represents your code
    -   **Client ID:** Codename
    -   **Client secret:** Client "password", but often not very secret
-   **Authorization server (auth):** Thing granting permission (usually the API owner)
    -   Often multiple servers (auth vs auth2 on next slide)
-   **Resource server (API):** Thing user is using (the API)

## The OAuth "dance" {-}

-   User to client: Hit this API for me!
-   Client (ID) to auth: Can I act as this user and do these things?
-   Auth to user: Is this ok?
-   User to auth: Yes
-   Auth to client's home address: Use this to get a key
-   Client (ID + secret) to auth2: Turn this into a key (I'm really me!)
-   Auth2 to client: Here's your key (and I'll log what it can do)
-   Client to API: Here's my key
-   API to auth: Can this key do this?
-   Auth to API: (checks logged scopes) Yes!
-   API to client: Ok, here's the info!

TODO: Image of OAuth dance

## OAuth credential dangers {-}

-   游릭 Client id: Like knowing a package name.
-   游리 Client secret:
    -   Can pretend to be your client, but user still needs to say ok
    -   Can your **client** do anything special?
        -   Installed (e.g. Slack app)?
        -   API usage limits?
-   游리 Authorization code: Unlikely to be an issue
    -   Only sent to provided redirect_uri
    -   Extremely short lived (often minutes or less)

## OAuth credential dangers (cont) {-}

-   游리 Refresh token: A longer-lived authorization code
    -   Can be used to get a new access token
    -   Usually revoked if you auth from scratch
    -   Also need client secret to refresh
-   游 Access token: The thing we're protecting
    -   Can do whatever it's authorized to do
    -   Usually easy to revoke
-   游댮 Username + password: We don't want to know these

## httr2::oauth_client() {-}

-   Almost definitely only need these (from API provider):
    -   `id` = Client ID
    -   `token_url` = URL where clients exchange authorization codes for tokens
    -   `secret` = Client secret
    -   Often `auth = "header"`
    -   If this client has multiple uses: `name` = unique for this use case
-   Construct once & reuse

## OAuth client demo {-}

```{r auth-oauth-youtube-client, eval = FALSE}
library(httr2)
yt_client <- oauth_client(
  Sys.getenv("YOUTUBE_CLIENT_ID"), 
  "https://oauth2.googleapis.com/token",
  secret = Sys.getenv("YOUTUBE_CLIENT_SECRET")
)
```

## httr2::req_oauth_auth_code() {-}

-   `auth_url` = URL to get an authorization code (from API)
-   `scope` = Usually comma-separated string of permissions
-   `pkce` = Good if supported, often have to turn off
-   `redirect_uri` = Where to send response
    -   Often need specific local port "http://localhost:4242" or "http://127.0.0.1:4242"
    -   Must be configured as part of client configuration (at API)
-   `cache_disk` = Set this `TRUE` if you can
    -   `cache_key` if you'll use this client for multiple tokens

## Oauth request demo {-}

```{r auth-oauth-youtube-call, eval = FALSE}
playlists <- request("https://youtube.googleapis.com/youtube/v3") |> 
  req_url_path_append("playlists") |> 
  req_url_query(part = "snippet", mine = TRUE, maxResults = 50) |> 
  req_oauth_auth_code(
    yt_client, 
    "https://accounts.google.com/o/oauth2/v2/auth",
    scope = "https://www.googleapis.com/auth/youtube",
    redirect_uri = "http://127.0.0.1:8888"
  ) |> 
  req_perform()
```

## Automating OAuth {-}

-   *If you can,* use httr2 cache: easiest, but
    -   auto-deletes when 30 days old
    -   fills logs w/ "Caching httr2 token in ..." messages
-   `httr2::req_oauth_bearer_jwt()` if you have JSON web token (service account)
-   `httr2::req_oauth_refresh()` if you have a refresh token
    -   `httr2::oauth_flow_auth_code()` once to get refresh

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
