# How can I test my plumber API?

**Learning objectives:**

-   Minimize API-specific code in your functions.
-   Wrap an API into an R package.
-   Test functions with {testthat}.
-   Test API-specific functions with `req` arguments.
-   Test {plumber} routers.

Partially inpsired by [API as a Package series from Jumping Rivers](https://www.jumpingrivers.com/blog/api-as-a-package-testing/)

## Pure Functions {-}

-   No hidden inputs
-   No side effects
-   Given inputs, always return same output.

*Abstract pure functions out of endpoint functions for easier testing.*

## Example: EXAMPLE NAME {-}

TODO: Create a simple API. First show as one function, then abstract out pure functions.

## plumber APIs as packages {-}

-   [R Packages](https://r-pkgs.org/) book for intro to R packages
-   Functions in `R/`
-   APIs in `inst/plumber/API_NAME/`
    -   Recommended: Use `entrypoint.R` to build router programmatically
-   Launch with `plumber::plumb_api("yourpkg", "API_NAME")`

## Brief Intro to testthat {-}

Much more in [R Packages](https://r-pkgs.org/)!

-   `usethis::use_testthat()` in package
-   Tests in `tests/testthat/test-filename.R`
-   `test_that("What you expected", { code to test that })`
-   Pure functions are relatively easy to test!

## Example: EXAMPLE FROM ABOVE IN A PACKAGE {-}

TODO: Fill this in, describing where the bits from above go inside a package.

## req helper {-}

```{r plumber-test-as-mock-req, eval = FALSE}
# tests/testthat/helper-req.R
as_mock_req <- function(...,
                        body = list(), 
                        path = list(), 
                        query = list(),
                        cookies = list()) {
  list2env(c(
    list(...), 
    argsBody = body, argsPath = path, argsQuery = query,  cookies = cookies
  ))
}
```

## Testing API-specific functions {-}

TODO: FILL THIS IN WITH SPECIFICS

```{r plumber-test-api-function, eval = FALSE}
test_that("THE THING WORKS", {
  req <- as_mock_req(body = list(whatever = 1))
  expect_equal(function_to_test(req), expected_result)
})
```

## local_api() {-}

```{r plumber-test-pr-run-bg, eval = FALSE}
# tests/testthat/helper-plumber.R
pr_run_bg <- function(pr, port) {
  callr::r_bg(
    plumber::pr_run,
    list(pr = pr, port = port),
    package = "plumber"
  )
}
local_api <- function(pr, port, env = parent.frame()) {
  bg_api <- pr_run_bg(pr, port)
  # Sys.sleep(2) # TODO: Test whether this is necessary!
  withr::defer(bg_api$kill(), envir = env)
}
req_local_api <- function(endpoint, port) {
  httr2::request(glue::glue("http://127.0.0.1:{port}")) |> 
    httr2::req_url_path_append(endpoint)
}
```

## Testing plumber routers {-}

Use these for auth, errors, etc

```{r plumber-test-router, eval = FALSE}
# test/testthat/test-router.R
test_that("My router does an expected thing", {
  api <- plumber::plumb_api("mypackage", "target_api")
  port <- httpuv::randomPort()
  local_api(api, port)
  # TODO: Update this with auth failure.
  req <- req_local_api("predict/x", port)
  expect_error(
    httr2::req_perform(req),
    class = "mypkg_error_auth"
  )
})
```

## random notes {-}

Include ideas from these [plumber](https://www.rplumber.io/) articles:

-   [Runtime](https://www.rplumber.io/articles/execution-model.html)

[JumpingRivers had a blog about this!](https://www.jumpingrivers.com/blog/api-as-a-package-testing/)

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
