# How can I handle API errors?

**Learning objectives:**

-   Describe how {plumber} handles errors by default.
-   Send different error messages to the server than to the client.
-   Describe the possible responses from an endpoint.
-   Return useful error responses from an endpoint.

## Plumber's default error handling {-}

```{r plumber-errors-default, eval = FALSE}
#* @get /simple
function(){
  stop("I'm an error!")
}

#> {"error":["500 - Internal server error"],"message":["Error in (function () : 
#> I'm an error!\n"]}
```

-   HTTP status code 500 for all errors
-   Error sent to client
-   Error also printed in terminal

## Custom error handling: basics {-}

```{r plumber-errors-set-simple, eval = FALSE}
my_error_handler <- function(req, res, err) {
  res$status <- 500 # Or something else, see below.
  
  # Do something server-side (logs)
  print(err)
  
  # Return an error JSON object.
  list(error = "An error occurred. Please contact your administrator.")
}

pr("plumber.R") |> 
  pr_set_error(my_error_handler) |> 
  pr_run()
```

## Logging strategies {-}

-   `print()` for function output
-   `cat()` for messages
-   `cli::cli_inform()` for formatted messages
-   `{log4r}` package for neater logs
-   See [DevOps for Data Science: Logging and Monitoring](https://do4ds.com/chapters/sec1/1-4-monitor-log.html) for more

## Error responses {-}

***API errors are still responses***

-   Status code specifies broad error category (see next slide)
    -   `400` to `499` = Client errors (user did something wrong)
    -   `500` to `599` = Server errors (something is broken)
    -   Reminder: By default, {plumber} uses `500`
-   Best practice: Error format same as successful response
    -   Be careful about this, don't make silent errors easy!
-   Need `plumber::pr_handle()` family for advanced options

## HTTP error status codes {-}

|code|title                |usage|
|:---|:--------------------|:----|
|400 |Bad Request          |Missing or bad argument, etc|
|401 |Unauthorized         |No or invalid authentication|
|403 |Forbidden            |Known user can't do that|
|404 |Not Found            |`path` arg doesn't exist|
|502 |Bad Gateway          |Your api hits another api, got error|
|504 |Gateway Timeout      |Your api hits another api, timed out|
|500 |Internal Server Error|Catch-all for other errors|

See [MDN HTTP response status codes](https://developer.mozilla.org/en-US/docs/Web/HTTP/Status#client_error_responses) for more options and further details

## Error descriptions in plumber block {-}

```{r plumber-output-response-annotation, eval = FALSE}
#* @response 400 Bad request. This usually occurs because of a missing or 
#*   malformed parameter.
#* @response 401 Unauthorized. A valid API token was not provided.
#* @response 403 Forbidden. This user cannot access that file.
```

-   `@response` `Status Code` `Description`
-   Can *only* give description (not object details)
-   This is a *promise,* but isn't *validated* by {plumber}

## Recommended error_handler {-}

```{r plumber-errors-error-handler, eval = FALSE}
# TODO: Finish working this out!
error_handler <- function(req, res, err) {
  res$status <- attr(err, "status_code") %||% 500L
  
  # TODO: Better understanding of log4r or another option.
  do_the_log_thing(err)

  # Skip serialization to always return JSON  
  res$headers <- list("Content-type" = "application/json")
  res$body <- jsonlite::toJSON(unclass(err))
  return(res) 
}
```

## Programmatic responses {-}

TODO: Adapt/finish this slide. And probably add one more about the internal error handling in plumber.R.

-   Can specify full OpenAPI Response Object programmatically
-   Impacts auto-generated Swagger documentation
-   Not well documented (outside of this book)

```{r plumber-output-response-programmatic, eval = FALSE}
pr_get(
  "/pathToEndpoint", function(res, req) { ... },
  responses = list(
    "403" = list(
      description = "Forbidden",
      content = list("application/json" = list())
    )
  )
)
```


## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
