# How can I get inputs from API users?

**Learning objectives:**

-   Decide whether API inputs should be passed in the path, query, body, header, or cookie.
-   Process API arguments directly from the request object.
-   Process API arguments specified in the path.
-   Process API arguments passed in the query string.
-   Process API arguments passed in the request body.
-   Process API arguments sent as a header.
-   Process API arguments stored in a cookie.

## Path vs Other {-}

-   **Path** for main noun (the thing you're acting on)
-   **Query/body** for arguments
-   **Cookies/header** for metadata

## Types of inputs {-}

**How will your API be used?** 

-   Directly in browser (Are you sure?)
    -   `GET`
    -   **Query** for things user decides
    -   **Cookies** for state, etc
-   Mostly programmatic (including from Shiny, etc)
    -   `POST`, `PUT`, `PATCH`, `DELETE`
    -   **Body** for things related to this specific request
    -   **Header** for things that aren't function arguments
    -   Possibly **cookies** for state still

## API function arguments {-}

-   *Can* expect named arguments, BUT
-   Better: Accept `req` (request object)
-   Specify priority (or only accept certain ones)
    -   Good documentation MUCH better than flexibility!
    -   By default: `argsQuery` > `argsPath` > `argsBody`

```{r plumberio-req-args, eval = FALSE}
function(req) {
  x <- req$argsPath$x %||% req$argsBody$x %||% req$argsQuery$x %||% 
    req$cookies$x %||% req$HTTP_X
}
```

## Parameter types {-}

-   Optionally specify type (~class) in API definition
    -   `@param x:int` (more details later)
-   Enclose in `[]` to allow length > 1 (array)
    -   `@param x:[int]`

## Parameter types table {-}

|type     |R class  |in         |
|:-------:|:-------:|:---------:|
|"boolean"|logical  |query, path|
|"number" |numeric  |query, path|
|"integer"|integer  |query, path|
|"string" |character|query, path|
|"object" |list     |body       |
|"file"   |raw      |body       |


## Path parameters {-}

```{r plumberio-path, eval = FALSE}
#* @get /users/<id>
function(id) {...} # better: use req$argsPath$id
```

Optional: only route certain types

```{r plumberio-path2, eval = FALSE}
#* @get /users/<id:int>
function(id) {...} # better: use req$argsPath$id
```

{plumber} casts typed *path* input

## Query parameters {-}

`https://example.com/api/repeat?letter=x&number=1`

```{r plumberio-query, eval = FALSE}
#* @get repeat
#* @param letter:string The letter.
#* @param number:integer The number of times to repeat it.
```

{plumber} does *not* cast *query* input!

## Request bodies {-}

-   `req$body` ≈ `req$bodyArgs`
    -   `body` sometimes has raw `value` + `parsed`, other details
-   Parsers can prepare body
    -   Specify as `#* @parser PARSER` (`json`, `csv`, `rds`, etc)
    -   Args in list: `#* @parser json list(simplifyVector = FALSE)`

```{r plumberio-body-parser, eval = FALSE}
#* @param df:object The input data.frame. Must contain columns "first_name" 
#*        and "city".
#* @parser feather list(col_select = c("first_name", "city"))
```

## Stabilizing input {-}

TODO: {stbl} package introduction once it's stable.

## HTTP headers {-}

Headers are processed and included in `req` object:

-   `-` replaced with `_`
-   UPPERCASE
-   `HTTP_` prefix

`X-custom-header` → `req$HTTP_X_CUSTOM_HEADER`

## Inputs in Cookies {-}

-   `req$cookies` = named list of (unencrypted) cookies
-   `req$session` = named list from encrypted cookies
    -   Much more info when we *save* cookies

## random notes {-}

Include ideas from these [plumber](https://www.rplumber.io/) articles:

-   [Routing & Input](https://www.rplumber.io/articles/routing-and-input.html)
    -   Specifically input bits. Also Dynammic Routes
-   [Rendering Output](https://www.rplumber.io/articles/rendering-output.html)

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
