---
engine: knitr
title: Quickly parse API responses
resources:
  - "schemas/demo_json.json"
---

# TODO

::: nonincremental
-   Merge back into "Get started with APIs"?
:::


# ️✅ Learning objectives

::: nonincremental
-   Parse nested lists with {tibblify}.
-   Use an API's OpenAPI description to determine the expected format of responses.
-   Parse API responses with {tibblify} and the response description.
:::

```{r}
#| label: schemas-packages-used
#| eval: true
#| warning: false
library(jsonlite)
library(tidyr)
library(dplyr)
library(tibblify)
library(yaml)
library(waldo)
```

## Aside: Why rectangle?

-   Apps usually "think" in objects
    -   Preferred data: Objects-with-properties ([example json](schemas/demo_json.json))
-   Data scientists usually compare many objects at once
    -   Preferred data: data frames
        -   Columns of variables (of same class)
        -   Rows of observations (~objects)
-   APIs are usually designed for the first model
-   Even data APIs tend to "think" in objects
    -   Because most programmers do

::: notes
-   This really belongs in the previous chapter or maybe intro, but I'm putting it here for this cohort.
-   Apps and thus most programmers.
    -   Even when you get a list of things, like on Amazon, each of those product blocks in the list is an html div object. It's rarely really columns of similar data.
-   Maybe this is obvious to some, but this felt like an epiphany to me.
-   Fundamental difference between R and everything else.
    -   Things like pandas/polars are attempts to make Python think like R.
-   Really pause to discuss.
:::

# The tibblify package

## Typical JSON data

```{r}
#| label: schemas-write_demo_json
#| include: false
#| eval: false
all_apis <- jsonlite::fromJSON(here::here("slides", "apis-start", "list.json"))
demo_json <- all_apis[c("apis.guru", "fec.gov", "googleapis.com:youtube")] |> 
  tibble::enframe(name = "api_id") |> 
  tidyr::unnest_wider(value) |> 
  tidyr::unnest_longer(versions, indices_to = "version") |>
  dplyr::filter(preferred == version) |>
  dplyr::select(api_id, version, versions) |> 
  tidyr::unnest_wider(versions) |> 
  dplyr::select(-info, -externalDocs) |> 
  # I want lists not a dataframe.
  dplyr::mutate(
    versions = purrr::pmap(
      list(version, added, updated, swaggerUrl, swaggerYamlUrl, openapiVer, link),
      \(version, added, updated, swaggerUrl, swaggerYamlUrl, openapiVer, link) {
        setNames(
          list(
            list(
              added = added, 
              updated = updated, 
              swaggerUrl = swaggerUrl, 
              swaggerYamlUrl = swaggerYamlUrl, 
              openapiVer = openapiVer, 
              link = link
            )
          ),
          version
        )
      }
    ),
    .keep = "unused"
  )
# Finally, I want a list, not a df.
demo_json2 <- purrr::map(
  demo_json$versions, \(version) list(versions = version)
)
names(demo_json2) <- demo_json$api_id
jsonlite::write_json(
  demo_json2, 
  here::here("slides", "schemas", "demo_json.json"), 
  auto_unbox = TRUE
)
```

[demo_json.json](schemas/demo_json.json)

```{r}
#| label: schemas-read_demo_json
#| include: false
#| eval: true
demo_json <- jsonlite::fromJSON(
  here::here("slides", "schemas", "demo_json.json")
)
```

::: fragment
```{r}
#| label: schemas-read_demo_json-url
url <- "https://dslc-io.github.io/club-wapir/slides/schemas/demo_json.json"
demo_json <- jsonlite::fromJSON(url)
```
:::

::: notes
-   When you click through, point out that the object-per-api is a perfectly reasonable standard, but that this data makes way more sense in R with a row per api, but a standardized column for each variable, like we saw in the last chapter & last slide
-   Note that sometimes `jsonlite::fromJSON()` mangles data, `jsonlite::read_json()` is same thing without auto-parsing.
:::

## Rectangling demo_json manually

```{r}
#| label: schemas-rectangle_demo_json-manual
#| eval: true
demo_json |> 
  tibble::enframe(name = "api_id") |>
  tidyr::unnest_longer(value, indices_include = FALSE) |> 
  tidyr::unnest_longer(value, indices_to = "version") |>
  tidyr::unnest_wider(value)
```

::: notes
-   Wouldn't it be nice if we could do all of that automatically?
:::

## Introducing tibblify

-   {tibblify} 📦 to auto-convert hierarchical data to tibbles
-   Super-charged `tidyr::unnest_auto()`
-   Tibbles all the way down
-   Experimental functionality for APIs

::: notes
-   (bullet 4): This probably belongs in a separate package, but for now it's in a dev branch of tibblify.
:::

## Rectangling demo_json with tibblify

::: fragment
```{r}
#| label: schemas-rectangle_demo_json-tibblify_auto
#| eval: true
dj_tibblified <- tibblify::tibblify(demo_json)
dj_tibblified
```
:::
::: fragment
```{r}
#| label: schemas-rectangle_demo_json-tibblify_auto-unnest
#| eval: true
dj_tibblified |> 
  dplyr::rename(api_id = ".names") |> 
  tidyr::unnest(versions)
```
:::

::: notes
-   At first tibblified looks similar to result of `enframe()`.
-   By default anything it has to name gets `.names`
    -   We'll see how to clean that up later.
-   But the nested things are tibbles, so we can use `tidyr::unnest()`
:::

## Rectangling manually vs tibblify

:::: columns
::: {.column .fragment}
```{r}
#| label: schemas-rectangle_demo_json-manual-side_by_side
#| eval: true
dj_tidyr <- 
  demo_json |> 
  tibble::enframe(name = "api_id") |>
  tidyr::unnest_longer(
    value, indices_include = FALSE
  ) |> 
  tidyr::unnest_longer(
    value, indices_to = "version"
  ) |>
  tidyr::unnest_wider(value)
```
:::
::: {.column .fragment}
```{r}
#| label: schemas-rectangle_demo_json-tibblify_auto-side_by_side
#| eval: true
dj_tibblify <- 
  demo_json |>
  tibblify::tibblify() |> 
  dplyr::rename(api_id = ".names") |> 
  tidyr::unnest(versions) |> 
  dplyr::rename(version = ".names")
```
:::
::::

<br>

::: fragment
```{r}
#| label: schemas-rectangle_demo_json-tibblify_auto-compare
#| eval: true
waldo::compare(
  dj_tibblify, dj_tidyr, 
  list_as_map = TRUE # Ignore column order
)
```
:::

::: notes
-   {waldo} is a package for finding differences.
-   `list_as_map = TRUE` means "ignore column order" here.
-   What did tibblify do to figure out how to rectangle the data?
:::

## tspec_guess()

-   `tibblify::tibblify()` `spec` argument
    -   "What should this look like?"
-   Guessed by default with `tibblify::guess_tspec()`

::: fragment
```{r}
#| label: schemas-tibblify-tspec_guess
#| eval: true
tibblify::guess_tspec(demo_json)
```
:::

::: notes
-   Point out `.names_to`.
-   Point out date columns ("added" and "updated").
-   It would be nice if we had a way to tell it what to expect.
:::

# The OpenAPI Specification

## Multiple Standards

-   Swagger 2.0 ➡️ OpenAPI 2.0
-   **OpenAPI 3.x**
-   OpenAPI 4.0 (in development)
-   Postman Collection
-   API Blueprint
-   Web Application Description Language (WADL)

::: notes
-   Swagger was created by wordnik.com, free online dictionary that compiles a bunch of sources (and wanted a way to store API info consistently).
-   Swagger specification & tools bought by SmartBear, spec donated to OpenAPI Initiative to make it officially open forever.
-   Swagger now technically means the tools, OpenAPI is the spec.
    -   But people talk about "swagger specs" all the time.
-   We'll focus on OpenAPI 3
-   We might talk about tools to convert others to OpenAPI 3.
:::

## YAML

:::: columns
::: {.column .fragment}
JSON <br>
(`jsonlite::read_json()`)
```{json}
{
  "info": {
    "contact": {
      "email": "mike.ralphson@gmail.com",
      "name": "APIs.guru",
      "url": "https://APIs.guru"
    },
    "title": "APIs.guru",
    "version": "2.2.0",
    "x-apisguru-categories": [
      "open_data",
      "developer_tools"
    ]
  },
    "security": []
}
```
:::
::: {.column .fragment}
YAML <br>
(`yaml::read_yaml()`)
```{yaml}
info:
  contact:
    email: mike.ralphson@gmail.com
    name: APIs.guru
    url: https://APIs.guru
  title: APIs.guru
  version: 2.2.0
  x-apisguru-categories:
    - open_data
    - developer_tools
security: [] # No security needed
```
:::
::::

::: notes
-   Originally "Yet Another Markup Language"
    -   Meaning of name changed because they wanted to stress it's for data, not documents per se.
    -   "YAML Ain't Markup Language"
-   YAML adds comments, readability
    -   Quoting optional for strings
    -   Indentation for children, rather than brackets
-   All JSON strings are valid YAML, not all YAML is valid JSON.
-   .yaml or .yml
-   You might have seen it:
    -   RMarkdown/Quarto headers & config
    -   Package configuration files (`_pkgdown.yml`, `codecov.yml`)
    -   GitHub Actions
:::

## Exploring an API Description

-   [APIs.guru](https://api.apis.guru/v2/specs/apis.guru/2.2.0/openapi.yaml)

::: fragment
```{yaml}
#| label: schemas-api_desc_yaml-specification
#| code-line-numbers: 1|2-5|6|7|8-10|11
openapi: 3.0.0 # Specification version number
info: # title, version, description, contact, license
servers: # One or more URLs + optional descriptions
tags: # Optional name, description, externalDocs of endpoint categories
externalDocs: # Optional url & description of additional documentation
security: # Optional list of named default security schemes
paths: # Endpoints of the API
webhooks: # Description of endpoints YOU can specify for API to SEND to
jsonSchemaDialect: # URI to extend components/schemas
x-whatever: # Extend with additional properties
components: # Reusable schemas, securitySchemes, reusable pieces of everything above
```
:::

::: notes
-   Technically the OpenAPI specification is the standard, and each API has a "description" that's defined in a "document" (which can be a composite of multiple files).
-   People can and will do things wrong/weird.
-   openapi = Spec version number, not API
-   Info, servers, tags, externalDocs = Info about the API.
-   security = default schemes to use, schemes are described in components/securitySchemes
-   paths = the things the API can do, detailed in later chapter(s)
-   webhooks, jsonSchemaDialect, extensions = we won't talk about these much if at all
-   Components is for storing reusable pieces of info about the API.
:::

## apis.guru APIs Schema

::: fragment
```{yaml}
#| label: schemas-apis_schema_yaml
#| code-line-numbers: 1-10|1-2|3|6-8|10|9|4-5|3-10
components:
  schemas:
    APIs:
      additionalProperties:
        $ref: "#/components/schemas/API"
      description: |
        List of API details.
        It is a JSON object with API IDs(`<provider>[:<service>]`) as keys.
      minProperties: 1
      type: object
```
:::
::: fragment
```{r}
tspec_apis <- tspec_df(
  .names_to = "api_id",
  tspec_api
)
```
:::

::: notes
-   The thing we're loading is described in APIs schema
-   I've simplified this slightly from the full openapi object.
-   JSON doesn't require a name for the names, so we need to pick something ("api_id")
-   additionalProperties --> tspec_, properties --> tib_
:::

## apis.guru API Schema

::: fragment
```{yaml}
#| label: schemas-api_schema_yaml
#| code-line-numbers: 1-15|3|5|15|4|13-14|6-12|12|10|11|8-9|3-15
components:
  schemas:
    API:
      additionalProperties: false
      description: Meta information about API
      properties:
        versions:
          additionalProperties:
            $ref: "#/components/schemas/ApiVersion"
          description: List of supported versions of the API
          minProperties: 1
          type: object
      required:
        - versions
      type: object
```
:::
::: fragment
```{r}
tspec_api <- tspec_row(
  tib_df(
    "versions", 
    .names_to = "version", 
    tspec_api_version
  )
)

```
:::

::: notes
-   `additionalProperties: false` means `tspec_row()`
-   `versions` property needs a `tib_` function.
-   `versions` is an `object`, so `tib_df` (or `tib_row`)
-   `versions` has `additionalProperties`, so defined in `tspec_df()`.
-   
:::

## apis.guru ApiVersion Schema

```{yaml}
#| label: schemas-api_version_schema_yaml
#| code-line-numbers: 1-35|3|35|4|29-34|5-28
components:
  schemas:
    ApiVersion:
      additionalProperties: false
      properties:
        added:
          description: Timestamp when the version was added
          format: date-time
          type: string
        link:
          description: Link to the individual API entry for this API
          format: url
          type: string
        openapiVer:
          description: The value of the `openapi` or `swagger` property of the source definition
          type: string
        swaggerUrl:
          description: URL to OpenAPI definition in JSON format
          format: url
          type: string
        swaggerYamlUrl:
          description: URL to OpenAPI definition in YAML format
          format: url
          type: string
        updated:
          description: Timestamp when the version was updated
          format: date-time
          type: string
      required:
        - added
        - updated
        - swaggerUrl
        - swaggerYamlUrl
        - openapiVer
      type: object
```

## apis.guru ApiVersion tspec

```{r}
#| eval: true
#| code-line-numbers: 1-10|11-18|12-13|14-16|17
tib_chr_datetime <- function(key, ..., required = TRUE) {
  tibblify::tib_scalar(
    key = key,
    ptype = vctrs::new_datetime(tzone = "UTC"),
    required = required,
    ptype_inner = character(),
    transform = \(x) as.POSIXct(x, format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC"),
    ...
  )
}
tspec_api_version <- tspec_row(
  tib_chr_datetime("added"),
  tib_chr_datetime("updated"),
  tib_chr("openapiVer"),
  tib_chr("swaggerUrl"),
  tib_chr("swaggerYamlUrl"),
  tib_chr("link", required = FALSE)
)
```
```{r}
#| include: false
#| eval: true
tspec_api <- tspec_row(
  tib_df("versions", tspec_api_version, .names_to = "version")
)
tspec_apis <- tspec_df(
  .names_to = "api_id",
  tspec_api
)
```

## Using tspecs: version

```{r}
#| eval: true
tibblify(demo_json$apis.guru$versions$`2.2.0`, tspec_api_version)
```


## Using tspecs: api

```{r}
#| eval: true
tibblify(demo_json$apis.guru, tspec_api) |> tidyr::unnest(versions)
```

## Using tspecs: apis

```{r}
#| eval: true
tibblify(demo_json, tspec_apis) |> tidyr::unnest(versions)
```

{{< include zzz.txt >}}
