---
engine: knitr
title: How can I use an API description?
---

# ️✅ Learning objectives

::: nonincremental
-   Use an API's OpenAPI description to learn about the API.
-   Parse nested lists with {tibblify}.
-   Parse API responses with {tibblify} and the response description.
:::

```{r}
#| label: openapi-packages-used
#| eval: true
```

# API Descriptions

## API Descriptions

-   [APIs.guru](https://api.apis.guru/v2/specs/apis.guru/2.2.0/openapi.json)
-   [OpenFEC](https://api.apis.guru/v2/specs/fec.gov/1.0/openapi.json)
-   [YouTube Data API v3](https://api.apis.guru/v2/specs/googleapis.com/youtube/v3/openapi.json)
-   [Asana](https://api.apis.guru/v2/specs/asana.com/1.0/openapi.json)

::: notes
-   Technically the OpenAPI specification is the standard, and each API has a "description"
-   OpenAPI 3.0+ easier to use than Swagger (aka 2.0)
-   Still some confusing bits, such as `security` vs `components$securitySchemas`
-   People can and will do things wrong/weird.
:::

## Read an API description

```{r}
#| label: apis_start-api_spec-download
#| eval: false
#| include: false
download.file(
  "https://api.apis.guru/v2/specs/googleapis.com/youtube/v3/openapi.yaml",
  here::here("slides", "httr2", "apis-start", "youtube.yaml")
)
```
```{r}
#| label: apis_start-api_spec-load
#| eval: true
#| include: false
youtube_spec <- yaml::read_yaml(here::here("slides", "httr2", "apis-start", "youtube.yaml"))
```

::: fragment
```{r}
#| label: apis_start-api_spec
youtube_spec <- yaml::read_yaml("https://api.apis.guru/v2/specs/googleapis.com/youtube/v3/openapi.yaml")
```
:::
::: fragment
```{r}
#| label: apis_start-api_spec-examine
#| eval: true
names(youtube_spec)
```
:::

## Introduction to tibblify

```{r}
#| label: apis_start-api_spec-tibblify
#| eval: true
youtube_paths <- tibblify::tibblify(youtube_spec$paths) |> 
  dplyr::glimpse()
```

## OpenAPI References

```{r}
#| label: apis_start-api_spec-tibblify-ref
#| eval: true
youtube_paths$parameters[[1]]
```

## parse_openapi_spec()

```{r}
#| label: apis_start-api_spec-tibblify-openapi
#| eval: true
youtube_paths <- tibblify::parse_openapi_spec(youtube_spec)
```
::: fragment
```{r}
#| label: apis_start-api_spec-tibblify2
#| eval: true
youtube_paths
```
:::

::: notes
-   We're looking at a dev version of tibblify which I plan to push through
    -   `pak::pak("mgirlich/tibblify#191")`
-   "endpoint" is another term for "path" -- a url in the API
-   Deals with references from paths to components > schemas
-   Captions has 4 different methods, so its tibble has 4 rows.
:::

## Unnesting paths

```{r}
#| label: apis_start-api_spec-tibblify3
#| eval: true
youtube_paths |> tidyr::unnest(operations)
```

::: notes
-   `operation` aka method, we'll discuss these more later; GET means you can load it like a web page
-   Note rows 3-6 are that `captions` endpoint, and then 7 is `captions/{id}`
:::

## Path details

```{r}
#| label: apis_start-api_spec-tibblify4
#| eval: true
youtube_paths |> tidyr::unnest(operations) |> dplyr::glimpse()
```

::: notes
-   (`endpoint` = path)
-   (`operation` = `method` = more details in a few chapters)
-   `summary` is meant for a very brief description, not often used
-   `description` can have markdown
-   `operation_id` = unique string for this operation, meant for tools to use as a function name
-   `tags` = Categories; can have more than one per endpoint, but often just one
-   `parameters` = tibble of inputs
-   `request_body` = tibble of a fancier type of input (later chapter)
-   `responses` = tibble of expected output
-   `deprecated` = is this endpoint no longer intended to be used
-   `global_parameters` = tibble of shared inputs
-   This tibble contains MOST of the useful information about the API
    -   We'll dig into this tibble for a simple API in the next chapter
:::
