---
engine: knitr
title: 7. Scrape data from web pages 
---

# ️✅ Learning objectives

::: nonincremental
-   Decide whether to scrape data from a web page.
-   Use {polite} to responsibly scrape web pages.
-   Scrape complex data structures from web pages.
-   Scrape content that requires interaction.
:::

```{r}
#| label: rvest-packages-used
#| eval: true
library(polite)
library(rvest)
library(xml2)
library(chromote)
```

::: notes
-   "Scraping" a web page means programmatically processing the web page to extract data in a usable format.
    -   Silly personal pet peeve: "scraping" vs "scrapping".
-   Can be a single page or several.
-   Point out that everything after the very beginning is the same as parsing HTML responses from APIs, because web pages are just GET requests
:::

# Decide whether to scrape data

## Do I need to scrape this data?

-   Look for an API!
-   Try {[datapasta](https://cran.r-project.org/package=datapasta)} 📦
    -   RStudio Addins
-   If it's one time & over-complicated, consider other copy/paste strategies
-   Only scrape pages you need

::: notes
-   More on looking for APIs in chapter 9 "Find APIs"
-   We'll see one place to look for an API on the next slide.
-   Even consider manual copy/paste one field at a time if it's not a TON of data.
-   Point of that last bullet: Don't hit the server more than you need to
:::

## Can I legally scrape this data?

-   ⚠️ Legal disclaimers (but may be over-protective)
-   USA:
    -   Can't copyright facts,
    -   *CAN* copyright collections of facts in some cases (creative)
    -   [TidyTuesday 2023-08-29: Fair Use](https://tidytues.day/2023/2023-08-29)
-   Other places:
    -   Sometimes stricter ([EU](https://eur-lex.europa.eu/LexUriServ/LexUriServ.do?uri=CELEX:31996L0009:EN:HTML))
    -   Sometimes more lax
-   ✅ Personal use or nonprofit education usually ok
-   ⚠️ Personally Identifiable Information (PII)

::: notes
-   If it's important, consult a lawyer.
-   This is why recipe blogs have stories about the recipes.
-   That TidyTuesday dataset exists because I was writing these notes.
-   EU provides more protections for database creators and for user-generated data
-   PII: Or even user-generated information in some cases
-   User-generated example: Stack overflow content requires attribution.
:::

## Should I scrape this data?

`robots.txt`: [github](https://github.com/robots.txt), [wikipedia](https://en.wikipedia.org/robots.txt)

-   `User-agent: *` = everybody
-   Search for name(s) of package(s) 
-   Search for specific pages within site
-   Check root of site (`/`) and your particular subfolder 
-   These aren't (necessarily) legally binding
-   [{robotstxt}](https://docs.ropensci.org/robotstxt/) 📦 for parsing `robots.txt`
-   [{polite}](https://dmi3kno.github.io/polite/) 📦 wraps and applies {robotstxt}

::: notes
-   Check comments at the top for general feel (comments have `#` at start)
-   Both examples point to the API. This is the right way to write a robots.txt!
-   `robots.txt` might be OVER protective, but watch for notes on how to use it
-   Don't hit a site you like a billion times in fast succession, you could "use up" their monthly bandwidth, for example
-   Your actions COULD cost them money, so don't be a jerk!
-   We'll see the {polite} package shortly, which wraps {robotstxt}
:::

# Scrape non-tabular data

## Motivating example: Cheese

![[Castelmagno cheese from cheese.com](https://www.cheese.com/castelmagno/)](rvest/castelmagno_cheese.png)

::: notes
-   For the book, I'll host some data so we can be sure it won't change, but I'm using this example because it's roughly the shape I'm looking for.
-   "Coss-tell-man-yo"
-   If you look at the "cleaning script" in this week's TidyTuesday... it's not perfect. I've improved it since then.
:::

## Three steps of web scraping with {rvest}

See [R4DS Chapter 24: "Web Scraping"](https://r4ds.hadley.nz/webscraping) for a full introduction

1.  Load the page.
2.  Find the object(s) (observations & variables) you want.
3.  Extract variables (text) into R object(s).

::: notes
-   Step 1 = scraping, 2-3 also apply to API requests that return HTML or XML.
-   Pages = text, regardless of what they contain.
    -   Text might point to URL of image, but page = text.
-   Specific table? Specific blocks of info? 
    -   List of products
-   Even numbers = text until parsed.
:::

## Use {polite} to scrape respectfully

::: fragment
```{r}
#| label: rvest-polite-bow
polite::bow(
  url,                             # Page to scrape (or root)
  user_agent = "polite R package", # If you don't change this, min delay = 5
  delay = 5,                       # Seconds between requests
  times = 3,                       # Retries
  force = FALSE,                   # Clears memoised functions
  verbose = FALSE,                 # Useful to know why it failed
  ...
)
```
:::
::: fragment
```{r}
#| label: rvest-polite-scrape
polite::scrape(
  bow,              # Session opened with bow()
  query = NULL,     # Named list to add after `?` in URL
  accept = "html",  # Specify html, json, xml, csv, txt, etc
  content = NULL,   # Optional MIME type
  verbose = FALSE   # Useful to know why it failed
)
```
:::

::: notes
-   Polite is a great package, but docs are light.
    -   I'm resisting urge to rewrite it in httr2 with better documentation.
-   bow():
    -   Good idea to set a `user agent`. If you leave default, it *has* to be slow. Searches for "polite" or polite package author in your user agent to set slow rule.
    -   I'd set `delay` to 0; user agent requirements will override this.
    -   Ok to leave `times` as-is, probably.
    -   `force`: "Memoise" = if inputs are the same, use the previous result.
        -   Ie, don't hit the API again if we already have the result.
    -   I'd recommend setting `verbose` to TRUE unless you're doing a mostly safe, automatic scrape.
-   scrape()
    -   `query` is useful for updating target in a loop, we'll see that later
    -   `accept`: For scraping, you likely want HTML
    -   Leave `content` NULL probably
    -   Set `verbose` to TRUE while testing.
-   We'll see a third function later: `nod()`
:::

## Load the page: bow() + scrape()

::: fragment
```{r}
#| label: rvest-polite-bow-usage
session <- polite::bow(
  "https://www.cheese.com/castelmagno/",
  user_agent = "rvest/1.0.4 (Jon Harmon; mailto:jonthegeek+useragent@gmail.com)",
  delay = 0,
  verbose = TRUE
)
session
#> <polite session> https://www.cheese.com/castelmagno/
#>     User-agent: rvest/1.0.4 (Jon Harmon; https://wapir.io; mailto:jonthegeek+useragent@gmail.com)
#>     robots.txt: 0 rules are defined for 1 bots
#>    Crawl delay: 0 sec
#>   The path is scrapable for this user-agent
```
:::
::: fragment
```{r}
#| label: rvest-polite-scrape-usage
castelmagno_page <- polite::scrape(session)
```
:::

## Find the object(s)

[https://www.cheese.com/castelmagno](https://www.cheese.com/castelmagno)

-   [Select a specific element](https://r4ds.github.io/bookclub-r4ds/select-a-specific-element.html) from R4DS slides

::: notes
-   Load the page and inspect element
-   Whole container: class: "summary-points" so `.summary-points`
    -   Technically could use "ul.summary_points"
:::

## Selected elements

-   [html_element() vs html_elements()](https://r4ds.github.io/bookclub-r4ds/select-finer-grained-elements.html) from R4DS slides

::: fragment
```{r}
#| label: rvest-summary_points
summary_points <- castelmagno_page |> 
  rvest::html_elements(".summary-points li")
summary_points
#> {xml_nodeset (15)}
#> [1] <li class="summary_milk">\n               ...
#> [2] <li class="summary_country">\n            ...
#> [3] <li class="summary_region">\n             ...
#> [4] <li class="summary_family">\n             ...
#> [5] <li class="summary_moisture_and_type">\n  ...
#> [6] <li class="summary_fat">\n                ...
#> [7] <li class="summary_calcium">\n            ...
#> [8] <li class="summary_texture">\n            ...
#> [9] <li class="summary_rind">\n               ...
#> [10] <li class="summary_tint">\n              ...
#> [11] <li class="summary_taste">\n             ...
#> [12] <li class="summary_smell">\n             ...
#> [13] <li class="summary_vegetarian">\n        ...
#> [14] <li class="summary_vegan">\n             ...
#> [15] <li class="summary_alt_spelling">\n      ...
```
:::

::: notes
-   We really only have a single observation and separate variables here, so that rule of thumb doesn't always hold!
:::

## Extract variables

::: fragment
```{r}
#| label: rvest-summary_points-html_text2
cheese_variables <- summary_points |> 
  rvest::html_text2() |> 
  stringr::str_remove("^[^:]*: ") # "Remove anything up to ": "
```
:::
::: fragment
```{r}
#| label: rvest-summary_points-html_attr
names(cheese_variables) <- summary_points |> 
  rvest::html_attr("class") |> 
  stringr::str_remove("^summary_")
```
:::
::: fragment
```{r}
#| label: rvest-summary_points-cheese_data
cheese_data <- tibble::tibble(!!!cheese_variables)
#> $ milk              <chr> "Made from pasteurized or unpasteurized cow's, goat's and sheep's milk"
#> $ country           <chr> "Italy"
#> $ region            <chr> "Piedmont"
#> $ family            <chr> "Blue"
#> $ moisture_and_type <chr> "semi-hard"
#> $ fat               <chr> "34.2 g/100g"
#> $ calcium           <chr> "4768 mg/100g"
#> $ texture           <chr> "crumbly, dense and grainy"
#> $ rind              <chr> "washed"
#> $ tint              <chr> "ivory"
#> $ taste             <chr> "sharp, spicy, strong"
#> $ smell             <chr> "strong"
#> $ vegetarian        <chr> "no"
#> $ vegan             <chr> "no"
#> $ alt_spelling      <chr> "Castelmagno PDO, Castelmagno di alpeggio, Castelmagno prodotto della montagna"
```
:::

::: notes
-   html_text2() cleans up text, you probalby always want it.
-   html_attr() gets things inside the tag (tab to view of the page)
-   I like the "bang bang bang" method for named vector to tibble, there are other ways to do this, too.
-   I'll briefly show cleaning on the next page.
    -   Split (comma/and)-separated lists into list-vars
    -   Extract info from "milk"
    -   Clean up classes
:::

## Aside: Cleaning

```{r}
#| label: rvest-summary_points-cleaning
cheese_data |> 
  dplyr::mutate(
    vegetarian = vegetarian == "yes",
    vegan = vegan == "yes",
    dplyr::across(
      c("fat", "calcium"),
      \(x) as.double(stringr::str_remove(x, " m?g/100g"))
    ),
    dplyr::across(
      c(-"milk", -"vegetarian", -"vegan", -"fat", -"calcium"),
      \(x) stringr::str_split(x, "(, )|( and )")
    )
  ) |> 
  dplyr::mutate(
    pasteurized = stringr::str_detect(milk, "\\bpasteurized"),
    unpasteurized = stringr::str_detect(milk, "\\bunpasteurized"),
    animal = stringr::str_extract_all(milk, "(\\S+)(?='s)"),
    .before = "milk",
    .keep = "unused"
  ) |> 
  dplyr::glimpse()
#> Rows: 1
#> Columns: 17
#> $ pasteurized       <lgl> TRUE
#> $ unpasteurized     <lgl> TRUE
#> $ animal            <list> <"cow", "goat", "sheep">
#> $ country           <list> "Italy"
#> $ region            <list> "Piedmont"
#> $ family            <list> "Blue"
#> $ moisture_and_type <list> "semi-hard"
#> $ fat               <dbl> 34.2
#> $ calcium           <dbl> 4768
#> $ texture           <list> <"crumbly", "dense", "grainy">
#> $ rind              <list> "washed"
#> $ tint              <list> "ivory"
#> $ taste             <list> <"sharp", "spicy", "strong">
#> $ smell             <list> "strong"
#> $ vegetarian        <lgl> FALSE
#> $ vegan             <lgl> FALSE
#> $ alt_spelling      <list> <"Castelmagno PDO", "Castelmagno di alpeggio", "Castelmagno prodotto della montagna"…
```

::: notes
-   {readr} `parse_*()` functions can be handy for dates and numbers.
    -   My {stbl} package might eventually be better for web data.
:::

# Scrape interactive web pages

## read_html_live

-   Data: [HMdb.org](https://www.hmdb.org/geolists.asp)
    -   [robots.txt](https://www.hmdb.org/robots.txt)
-   Live-coding!

::: fragment
```{r}
#| label: rvest-read_html_live
session <- rvest::read_html_live("https://www.hmdb.org/geolists.asp")
session |> 
  rvest::html_element("#StatesList") |> 
  rvest::html_text2()
session |> 
  rvest::html_element("div.bodysansserif") |>
  rvest::html_elements("td:nth-child(2)") |> 
  rvest::html_text2()
session$click("tr:nth-child(1) .countryarrow")
session |> 
  rvest::html_element("#StatesList") |> 
  rvest::html_text2()
session$click("tr:nth-child(2) .countryarrow")
session |> 
  rvest::html_element("#StatesList") |> 
  rvest::html_text2()
session |> 
  rvest::html_element("#StateSidebar") |>
  rvest::html_elements("td:nth-child(2)") |> 
  rvest::html_text2()
session |> 
  rvest::html_element("#CountiesList") |> 
  rvest::html_text2()
session$click("#StateSidebar tr:nth-child(3) .statearrow")
session |> 
  rvest::html_element("#CountiesList") |> 
  rvest::html_text2()
session |> 
  rvest::html_element("#CountySidebar") |> 
  rvest::html_elements("a") |> 
  rvest::html_text2()
session |> 
  rvest::html_element("#CountySidebar") |> 
  rvest::html_elements("a") |> 
  rvest::html_attr("href")
```
:::
